### 2019-06-18 initial/notes.tex
add correct eta calc

Though this section is located before the weak tidal friction section, we
actually return to this second toy problem to get some practice on the Henrard
paper, to confirm the accuracy of our simulations compared to analytical
predictions, and to get some practice doing these new integrals. Consider the
dissipation-free Cassini state system with a small but constant $\pd{\eta}{t} =
\epsilon$. For clarity, the Hamiltonian is
\begin{equation}
    H = -\alpha\frac{\mu^2}{2} +
        \mu \cos I - \sin I \sqrt{1 - \mu^2} \theta \cos \phi.
\end{equation}
This proves to be the algebraically easiest formulation, instead of using
$\eta$.

@@

>>>>

Though this section is located before the weak tidal friction section, we
actually return to this second toy problem to get some practice on the Henrard
paper, to confirm the accuracy of our simulations compared to analytical
predictions, and to get some practice doing these new integrals. Consider the
dissipation-free Cassini state system with a small but constant $\pd{\eta}{t} =
\epsilon$. For clarity, the Hamiltonian is
\begin{equation}
    H = -\alpha\frac{\mu^2}{2} +
        \mu \cos I - \sin I \sqrt{1 - \mu^2} \cos \phi.
\end{equation}
This proves to be the algebraically easiest formulation, instead of using
$\eta$.

@@

\subsection{Retry without changing Hamiltonian}

Of course, it should be noted that modifying the Hamiltonian by dividing through
by $\eta$ and defining $\alpha = 1/\eta$ actually changes the dynamics; we can
see this since $\pd{(H/\eta)}{t}\dot{\eta} \neq
\pd{H}{t}\frac{\dot{\eta}}{\eta}$. Since my simulated system actually uses the
same Hamiltonian as toy problem 1, we should probably update the above algebra
to use the correct term. Note that the correction should be small, of order
$\frac{H/\eta^2}{\pd{H}{\eta}/\eta}$.

For clarity, $H$ and $H_4$ are now
\begin{align*}
    H &= -\frac{\mu^2}{2}
            + \eta\p*{\mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi}, &
    H_4 &= -\eta \sin I + \frac{\eta^2 \cos^2 I}{2}.
\end{align*}
Thus, our $B_i$ are
\begin{align*}
    B_i &= \int_{\mathcal{C}_i} \epsilon \mu \eta
            \s*{-\sin I + \eta \cos^2 I -
                \p*{\mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi}}
            \;\mathrm{d}t,\\
        &\approx \int_{\mathcal{C}_i} \epsilon \mu \eta
            \frac{\eta \cos^2 I - \mu \cos I
                + \sin I\p*{\cos \phi \sqrt{1 - \mu^2} - 1}}{-\mu + \eta \cos I}
            \;\mathrm{d}\phi,\\
        &= \int_{\mathcal{C}_i} \epsilon \mu \eta
            \p*{-\cos I + \sin I \frac{\cos \phi \sqrt{1 - \mu^2} - 1}{-\mu +
            \eta \cos I}} \;\mathrm{d}\phi.
\end{align*}
That the second term does not diverge requires some careful work:
\begin{align*}
    \frac{\cos \phi\sqrt{1 - \mu^2} - 1}{\mu - \eta \cos I}
        &\approx \frac{\sqrt{1 - 2\eta \sin I\p*{1 - \cos \phi}}\cos \phi
            - 1}{\mu - \eta \cos I},\\
        &\approx \frac{(1 - \eta \sin I(1 - \cos \phi))\cos \phi - 1}{
            \mu - \eta \cos I},\\
        &\approx \pm\frac{(1 + \eta \sin I)(\cos \phi - 1)}{\sqrt{2\eta \sin I
            (1 - \cos \phi)}},\\
        &\approx \mp\sqrt{\frac{1 - \cos \phi}{2\eta \sin I}},\\
    B_i &= \int_{\mathcal{C}_i} \epsilon \mu \eta
            \p*{-\cos I \pm \sin I
                \sqrt{\frac{1 - \cos \phi}{2\eta \sin I}}} \;\mathrm{d}\phi.
\end{align*}
Note that the plus sign corresponds to the top integral, where $\mu - \eta \cos
I > 0$. The top integral evaluates easily
\begin{align*}
    B_{top} &= \int\limits_{2\pi}^0 -\epsilon \mu \eta \cos I\;\mathrm{d}\phi
            + \frac{\epsilon}{2}\int\limits_{2\pi}^0
                \mu\sqrt{2\eta \sin I\p*{1 - \cos \phi}}\;\mathrm{d}\phi,\\
        &= \epsilon \eta \cos I\p*{2\pi \eta \cos I + 8\sqrt{\eta \sin I}}
            + \frac{\epsilon}{2}\p*{\eta \cos I (8\sqrt{\eta \sin I})
            + 4\pi \eta \sin I},\\
        &= \epsilon \eta \s*{
            2\pi \sin I + 12\sqrt{\eta \sin I}\cos I + 2\pi \eta \cos^2 I}.
\end{align*}
The bottom integral just flips the signs of the cross terms (the $\sqrt{\eta}$
terms), since the integration path changes, which changes the sign inside the
$B_i$ expression and $\mu$. The total integral is the difference between the top
and bottom integrals, which is just twice the sign-changed terms, and so we
obtain
\begin{equation}
    P_c = \frac{24\cos I \sqrt{\eta \sin I}}{
            2\pi \sin I + 12\sqrt{\eta \sin I}\cos I + 2\pi \eta \cos^2 I}.
\end{equation}
Indeed we get the same expression; this is probably since the approximations we
made in writing down $B_i$ washed out the extra $\eta$ dependence.
